---
title: "Corona virus in Ethiopia "
author: " Zenebe Ageru Yilma"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
  #social: ["facebook", "twitter", "linkedin"]
    source_code: embed
    vertical_layout: fill
---


```{r setup, include=FALSE }
#------------------ Packages ------------------#
library(flexdashboard)
library(tidyverse)
library(magrittr)
library(lubridate)
library(gridExtra)
library(kableExtra)
#install.packages("devtools")
#devtools::install_github("RamiKrispin/coronavirus")
#library(coronavirus)
#data(coronavirus)
#update_datasets()
# View(coronavirus)
#max(coronavirus$date)

`%>%` <- magrittr::`%>%`
#------------------ Parameters ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
confirmed_color <- "purple"
active_color <- "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "red"
#------------------ Data ------------------
urlconfirmed="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
data.confirmed <- read.csv(url(urlconfirmed), sep = "," )
urldeath="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
data.deaths <- read.csv(url(urldeath), sep = ",")
urlrecovered="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"
data.recovered <- read.csv(url(urlrecovered), sep = ",")
n.col <- ncol(data.confirmed)
data.confirmed %<>%pivot_longer(cols = 5:n.col, names_to = "date", values_to = "cases") %<>% mutate(date = date %>% substr(2,10) %>% mdy())%>%add_column(type = factor("confirmed"))
data.deaths %<>%pivot_longer(cols = 5:n.col, names_to = "date", values_to = "cases") %<>% mutate(date = date %>% substr(2,10) %>% mdy())%>%add_column(type = factor("death"))
data.recovered %<>%pivot_longer(cols = 5:n.col, names_to = "date", values_to = "cases") %<>% mutate(date = date %>% substr(2,10) %>%
                                                                                                   mdy())%>%add_column(type = factor("recovered"))

coronavirus <- rbind(data.confirmed, data.deaths, data.recovered)

coronavirus$date <-    as.Date(coronavirus$date,"%m/%d/%y")
coronavirus$Country.Region <- as.character(coronavirus$Country.Region)
coronavirus$Province.State<- as.character(coronavirus$Province.State)


df_w <- coronavirus %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::group_by(Country.Region, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(-confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(Country.Region == "United Arab Emirates", "UAE", Country.Region)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

dw_daily <- coronavirus %>%
  dplyr::group_by(date, type) %>%
  dplyr::summarise(total = sum(cases, na.rm = TRUE)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::arrange(date) %>%
  #dplyr::ungroup() %>%
  dplyr::mutate(active = confirmed - death - recovered) %>%
  dplyr::mutate(active = confirmed - death) %>%
  dplyr::mutate(
    confirmed_cum = cumsum(confirmed),
    death_cum = cumsum(death),
    recovered_cum = cumsum(recovered),
    active_cum = cumsum(active)
  )

df <- coronavirus %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::filter(Country.Region == "Ethiopia") %>%
  dplyr::group_by(Country.Region, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(-confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(Country.Region == "United Arab Emirates", "UAE", Country.Region)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

df_daily <- coronavirus %>%
  dplyr::filter(Country.Region == "Ethiopia") %>%
  dplyr::filter(date>="2020-03-01") %>%
  dplyr::group_by(date, type) %>%
  dplyr::summarise(total = sum(cases, na.rm = TRUE)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::arrange(date) %>%
  #dplyr::ungroup() %>%
  dplyr::mutate(active = confirmed - death - recovered) %>%
  dplyr::mutate(active = confirmed - death) %>%
  dplyr::mutate(
    confirmed_cum = cumsum(confirmed),
    death_cum = cumsum(death),
    recovered_cum = cumsum(recovered),
    active_cum = cumsum(active)
  )


df1 <- coronavirus %>% dplyr::filter(date == max(date))
```

In Ethiopia
=======================================================================

Row {data-width=400}
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total confirmed cases",
  icon = "fas fa-user-md",
  color = confirmed_color
)
```

### Recovered {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$recovered, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$recovered, na.rm = TRUE) / sum(df$recovered+df$death), 1),
    "%)",
    sep = ""
  ),
  caption = "Recovered cases (recovery rate)",
  icon = "fas fa-heart-broken",
  color = recovered_color
)
```




### death {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$death, na.rm = TRUE) / sum(df$death+df$recovered), 1),
    "%)",
    sep = ""
  ),
  caption = "Death cases (death rate)",
  icon = "fas fa-heart-broken",
  color = death_color
)
```

Row
-----------------------------------------------------------------------

### **Daily cumulative cases by type** (Ethiopia only)
    
```{r}
plotly::plot_ly(data = df_daily) %>%
  plotly::add_trace(
    x = ~date,
     #y = ~active_cum,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
     #name = "Active",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~ recovered_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Recovered",
    line = list(color = recovered_color),
    marker = list(color = recovered_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~ death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  )%>%
  plotly::add_annotations(
    x = as.Date("2020-3-13"),
    y = 1,
    text = paste("First case"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-4-5"),
    y = 1,
    text = paste("First death"),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-29"),
    y = 14,
    text = paste(
      "New containment",
      "",
      "measures"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  ) %>%
  plotly::layout(
    title = "",
    yaxis = list(title = "Cumulative number of cases"),
    xaxis = list(title = "Date"),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
    )
```



Global
=======================================================================

Row {data-width=400}
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df_w$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total confirmed cases",
  icon = "fas fa-user-md",
  color = confirmed_color
)
```

### Recovered {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df_w$recovered, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df_w$recovered, na.rm = TRUE) / sum(df_w$recovered +df_w$death), 1),
    "%)",
    sep = ""
  ),
  caption = "Recovered cases (recovery rate)",
  icon = "fas fa-heart-broken",
  color = recovered_color
)
```


### death {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df_w$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df_w$death, na.rm = TRUE) / sum(df_w$death+df_w$recovered), 1),
    "%)",
    sep = ""
  ),
  caption = "Death cases (death rate)",
  icon = "fas fa-heart-broken",
  color = death_color
)
```


Row
-----------------------------------------------------------------------

### **Daily cumulative cases by type** (Global)
    
```{r}
plotly::plot_ly(data = dw_daily) %>%
  plotly::add_trace(
    x = ~date,
     #y = ~active_cum,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
     #name = "Active",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~ recovered_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Recovered",
    line = list(color = recovered_color),
    marker = list(color = recovered_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~ death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  )%>%
  
  plotly::layout(
    title = "",
    yaxis = list(title = "Cumulative number of cases"),
    xaxis = list(title = "Date"),
    legend = list(x = 0.1, y = 0.9),
    hovermode = "compare"
  )
```


Comparison
=======================================================================


Top ten
-----------------------------------------------------------------------
### **Cases distribution top ten countries**

```{r}
df_top <- coronavirus %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::filter(Country.Region == "US" |
    Country.Region == "Spain" |
    Country.Region ==  "Italy"|
      Country.Region == "United Kingdom" |
      Country.Region =="Russia"  |
    Country.Region =="France" |
      Country.Region == "Germany"|
    Country.Region == "Brazil"|
      Country.Region == "Turkey"|
    Country.Region =="India" ) %>%
  dplyr::group_by(Country.Region, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(Country.Region == "United Arab Emirates", "UAE", Country.Region)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

plotly::plot_ly(
  data = df_top,
  x = ~country,
  # y = ~unrecovered,
  y = ~ confirmed,
  # text =  ~ confirmed,
  # textposition = 'auto',
  type = "bar",
  name = "Confirmed",
  marker = list(color = active_color)
) %>%
  plotly::add_trace(
    y = ~recovered,
    # text =  ~ recovered,
    # textposition = 'auto',
    name = "Recovered",
    marker = list(color = recovered_color)
  ) %>%
  plotly::add_trace(
    y = ~death,
    # text =  ~ death,
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  )%>%
  plotly::layout(
    barmode = "stack",
    yaxis = list(title = "Total cases"),
    xaxis = list(title = ""),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```




### **Cases distribution in Horn of Africa and Ethiopian Boarders**

```{r daily_summary}
df_hr <- coronavirus %>%
  dplyr::filter(date == max(date)) %>%
  dplyr::filter(Country.Region == "Ethiopia" |
    Country.Region == "Djibouti" |
    Country.Region == "Eritrea"|
      Country.Region == "South Sudan"|
      Country.Region == "Sudan"|
      Country.Region == "Kenya"|
    Country.Region == "Somalia") %>%
  dplyr::group_by(Country.Region, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(Country.Region == "United Arab Emirates", "UAE", Country.Region)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

plotly::plot_ly(
  data = df_hr,
  x = ~country,
  # y = ~unrecovered,
  y = ~ confirmed,
  # text =  ~ confirmed,
  # textposition = 'auto',
  type = "bar",
  name = "Confirmed",
  marker = list(color = active_color)
) %>%
  plotly::add_trace(
    y = ~death,
    # text =  ~ death,
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  ) %>%
  plotly::add_trace(
    y = ~recovered,
    # text =  ~ recovered,
    # textposition = 'auto',
    name = "Recovered",
    marker = list(color = recovered_color)
  ) %>%
  plotly::layout(
    barmode = "stack",
    yaxis = list(title = "Total cases"),
    xaxis = list(title = ""),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```

Daily New cases
=======================================================================

Row {data-width=400}
-------------------------------------


### ** top 5 Daily new cases **
    
```{r}
daily_confirmed <- coronavirus %>%
  dplyr::filter(type == "confirmed") %>%
  dplyr::mutate(country = Country.Region) %>%
  dplyr::group_by(date, country) %>%
  dplyr::summarise(total = sum(cases)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = country, values_from = total)

#----------------------------------------
# Plotting the data

daily_confirmed %>%
  plotly::plot_ly() %>%
  plotly::add_trace(
    x = ~date,
    y = ~US,
    type = "scatter",
    mode = "lines+markers",
    name = "USA"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Spain,
    type = "scatter",
    mode = "lines+markers",
    name = "Spain"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Brazil,
    type = "scatter",
    mode = "lines+markers",
    name = "Brazil"
  )  %>%
  plotly::add_trace(
    x = ~date,
    y = ~ `United Kingdom`,
    type = "scatter",
    mode = "lines+markers",
    name = "UK"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Russia,
    type = "scatter",
    mode = "lines+markers",
    name = "Russia"
  ) %>%
  plotly::layout(
    title = "",
    legend = list(x = 0.1, y = 0.9),
    yaxis = list(title = "Number of new cases"),
    xaxis = list(title = "Date"),
    # paper_bgcolor = "black",
    # plot_bgcolor = "black",
    # font = list(color = 'white'),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```

 
-------------------------------------


### **Daily new cases in the Horn of Africa**
    
```{r}
daily_confirmed <- coronavirus %>%
  dplyr::filter(type == "confirmed") %>%
  dplyr::filter(date >= "2020-03-10") %>%
  dplyr::mutate(country = Country.Region) %>%
  dplyr::group_by(date, country) %>%
  dplyr::summarise(total = sum(cases)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = country, values_from = total)

#----------------------------------------
# Plotting the data

daily_confirmed %>%
  plotly::plot_ly() %>%
  plotly::add_trace(
    x = ~date,
    y = ~Ethiopia,
    type = "scatter",
    mode = "lines+markers",
    name = "Ethiopia"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Djibouti,
    type = "scatter",
    mode = "lines+markers",
    name = "Djibouti"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Somalia,
    type = "scatter",
    mode = "lines+markers",
    name = "Somalia"
  )  %>%
  plotly::add_trace(
    x = ~date,
    y = ~Sudan,
    type = "scatter",
    mode = "lines+markers",
    name = "Sudan"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~`South Sudan`,
    type = "scatter",
    mode = "lines+markers",
    name = "South Sudan"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Kenya,
    type = "scatter",
    mode = "lines+markers",
    name = "Kenya"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Eritrea,
    type = "scatter",
    mode = "lines+markers",
    name = "Eritrea"
  ) %>%
  plotly::layout(
    title = "",
    legend = list(x = 0.1, y = 0.9),
    yaxis = list(title = "Number of new cases"),
    xaxis = list(title = "Date"),
    # paper_bgcolor = "black",
    # plot_bgcolor = "black",
    # font = list(color = 'white'),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```




Map
=======================================================================

### **World map of cases** (*use + and - icons to zoom in/out*)

```{r}
# map tab added by Art Steinmetz
library(leaflet)
library(leafpop)
library(purrr)
cv_data_for_plot <- coronavirus %>%
  # dplyr::filter(Country.Region == "Ethiopia") %>%
  dplyr::filter(cases > 0) %>%
  dplyr::group_by(Country.Region, Province.State, Lat, Long, type) %>%
  dplyr::summarise(cases = sum(cases)) %>%
  dplyr::mutate(log_cases = 2 * log(cases)) %>%
  dplyr::ungroup()
cv_data_for_plot.split <- cv_data_for_plot %>% split(cv_data_for_plot$type)
pal <- colorFactor(c("orange", "red", "green"), domain = c("confirmed", "death", "recovered"))
map_object <- leaflet() %>% addProviderTiles(providers$Stamen.Toner)
names(cv_data_for_plot.split) %>%
  purrr::walk(function(df) {
    map_object <<- map_object %>%
      addCircleMarkers(
        data = cv_data_for_plot.split[[df]],
        lng = ~Long, lat = ~Lat,
        #                 label=~as.character(cases),
        color = ~ pal(type),
        stroke = FALSE,
        fillOpacity = 0.8,
        radius = ~log_cases,
        popup = leafpop::popupTable(cv_data_for_plot.split[[df]],
          feature.id = FALSE,
          row.numbers = FALSE,
          zcol = c("type", "cases", "Country.Region", "Province.State")
        ),
        group = df,
           #            clusterOptions = markerClusterOptions(removeOutsideVisibleBounds = F),
        labelOptions = labelOptions(
          noHide = F,
          direction = "auto"
        )
      )
  })

map_object %>%
  addLayersControl(
    overlayGroups = names(cv_data_for_plot.split),
    options = layersControlOptions(collapsed = FALSE)
  )
```





About
=======================================================================

**The Coronavirus Dashboard: the case of Ethiopia**

This Coronavirus dashboard: the case of Ethiopia provides an overview of the 2019 Novel Coronavirus COVID-19 (2019-nCoV) epidemic for Belgium. This dashboard is built with R using the R Makrdown framework and was adapted from this [dashboard](https://ramikrispin.github.io/coronavirus_dashboard/){target="_blank"} by Rami Krispin.

**Code**

The code behind this dashboard is available on [GitHub](https://github.com/AntoineSoetewey/coronavirus_dashboard){target="_blank"}.

**Data**

The input data for this dashboard is the dataset available from the [`{coronavirus}`](https://github.com/RamiKrispin/coronavirus){target="_blank"} R package. Make sure to download the development version of the package to have the latest data:

```
install.packages("devtools")
devtools::install_github("RamiKrispin/coronavirus")
```

The data and dashboard are refreshed on a daily basis.

The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv){target="_blank"}.

**Contact**

For any question or feedback, follow on 
[Telegram](https://t.me/classicweb) |
[Facebook](https://www.facebook.com/zageru)| [Twitter](https://twitter.com/classichube)|
